Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
#library(ggplot2)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
library(ggplot2)
fb <- read.csv(file= 'pseudo_facebook.tsv',sep='\t')

table(fb$gender)

by(fb$friend_count, fb$gender, mean)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = fb, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

#print(fb$tenure)

#Output for Tenure as is

qplot(x = na.omit(fb$tenure), binwidth = 25,
      color = I('black'), fill = I('#00ff00'), xlab = 'Tenure in Years')+
  scale_x_continuous(breaks = 0:7)

#ggplot(data = fb) +
#  geom_histogram(aes(tenure),binwidth = 25)

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
library(ggthemes)
?seq()

#Output for tenure with NA and in Years

a <- na.omit(fb$tenure)

qplot(x = (a/365), binwidth = 0.25,
      color = I('black'), fill = I('#099DD9'), xlab = 'Tenure in Years')+
  scale_x_continuous(breaks = 0:7, limits = c(0,7))

ggplot(aes(x = tenure/365), data = fb) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')+
  scale_x_continuous(breaks = seq(0, 7, 0.5), limits = c(0,7))

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

qplot(x = age, data = fb, binwidth = 2,color = '#0000ff', fill = '#F79420')+
  scale_x_continuous(breaks = seq(0,120,5))

ggplot(aes(x = age), data = fb) +
  geom_histogram(binwidth = 1, color = 'black', fill = '#F79420')+
  scale_x_continuous(breaks = seq(0, 120, 5))

ggplot(aes(x = age), data = fb) +
  geom_bar()+
  scale_x_continuous(breaks = seq(0, 120, 5))
```

#### What do you notice?
Response:
Most people are of age 18, frequency is high in 20-25, and there are some unnatural spikes after 100 close to 93-94, this seems to be counter-intuitive.
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r transforming freind count}
print('Regular Summary')
summary(fb$friend_count)

print('log10 Summary')
summary(log10(fb$friend_count + 1))

print('SQRT Summary')
summary(sqrt(fb$friend_count))

qplot(x = friend_count, data = fb)+
  scale_x_log10()


```
***

### Add a Scaling Layer
Notes:

```{r fig.width=20, fig.height=30}
library(ggplot2)
library(gridExtra)
library(scales)
?trans_breaks
?scale_x_log10

p1 = ggplot(aes(x = (friend_count+1)), data = fb)+
  geom_histogram(binwidth = 0.05)+
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^(x)))+
  labs(x = 'Log(X) Friend Count')

p2 = ggplot(aes(x = (friend_count)), data = fb)+
  geom_histogram(binwidth = 0.5)+
  scale_x_sqrt(breaks = trans_breaks("sqrt", function(x) x ^ 2))+
  labs(x = 'sqrt(X) Friend Count')

p3 = ggplot(aes(x = (friend_count)), data = fb)+
  geom_histogram(binwidth = 50)+
  scale_x_continuous(breaks = seq(10,2500,250), limits = c(10,2500))+
  labs(x = '(X) Friend Count')

grid.arrange(p1, p2, p3, ncol=2)

```

***


### Frequency Polygons

```{r fig.width=10, fig.height=10}

library(ggthemes)

p4 = ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(fb, !is.na(gender)))+
  geom_freqpoly(aes(color = gender), binwidth = 10)+
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))+
  xlab('Friend Count')+
  theme_solarized()

p5 = ggplot(aes(x = friend_count, y = ..density..), data = subset(fb, !is.na(gender)))+
  geom_freqpoly(aes(color = gender), binwidth = 10)+
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))+
  xlab('Friend Count')+
  theme_solarized()

p6 = ggplot(aes(x = friend_count), data = subset(fb, !is.na(gender)))+
  geom_freqpoly(aes(color = gender), binwidth = 10)+
  scale_x_continuous(limits = c(0,1000), breaks = seq(0,1000,50))+
  xlab('Friend Count')+
  theme_solarized()

grid.arrange(p4, p5,p6, ncol = 1)

ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(fb, !is.na(gender)))+
  geom_freqpoly(aes(color = gender), binwidth = 10)+
  scale_x_continuous(limits = c(350,1000), breaks = seq(350,1000,20))+
  xlab('Friend Count')+
  theme_solarized()

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

library(data.table)

ggplot(aes(x = www_likes), data = subset(fb, !is.na(gender)))+
  geom_freqpoly(aes(color = gender), binwidth = 100)+
  scale_x_continuous(limits = c(0,350),breaks = seq(0,350,25))+
  xlab('Friend Count')+
  theme_solarized()

a <- fb$likes[between(fb$www_likes,1,8000)]
tail(a)
str(a)

df = c()
df2 = c()

i=1


for (x in fb$likes)
{ 
  if (x < 1) {}
    else {if (x < 800) {df <- c(df,x) }}
}

for (x in fb$likes)
{ 
  
  if (x < 1) {}
    else {if (x < 800) {df2 <- c(df2,fb$gender[i]) }}
  i=i+1
}

tail(df)
tail(df2)

write.csv(df, 'txt.csv')

by(fb$likes[(between(fb$www_likes,1, 800))],fb$gender[(between(fb$www_likes,1, 800))],summary)

summary(fb$likes > 0 & fb$likes < 80, fb$gender == 'female')
summary(fb$likes > 0, fb$gender == 'male')

?count

range(fb$likes[(fb$www_likes > 1 & fb$www_likes < 800)], na.rm = TRUE)

by(fb$likes,fb$gender,summary)

#limits = c(0,1000), breaks = seq(0,1000,50)

library(data.table)
sum(subset(fb$www_likes, fb$gender == 'male'))
sum(subset(fb$www_likes, fb$gender == 'female'))

```


***

### Box Plots
Notes:

```{r Box Plots}
library(ggplot2)

ggplot(aes(gender,friend_count), data= subset(fb,!is.na(gender)))+
  geom_boxplot()+
  scale_y_continuous(limits = c(1,750),
                     breaks = seq(0,750,100))

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
ggplot(aes(gender,friend_count), data= subset(fb,!is.na(gender)))+
  geom_boxplot()+
  scale_y_continuous(limits = c(0,1000),
                     breaks = seq(0,750,100))

ggplot(aes(gender,friend_count), data= subset(fb,!is.na(gender)))+
  geom_boxplot()+
  coord_cartesian(ylim =c(0,250))

#use cord_cartesian for zooming into data, without changing the axis.
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
ggplot(aes(gender,friendships_initiated), data= subset(fb,!is.na(gender)))+
  geom_boxplot()+
  coord_cartesian(ylim = c(1,200))

str(fb)

by(fb$friendships_initiated, fb$gender, mean)

dim(subset(fb,fb$gender == 'male'))

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
library(scales)

mobile_check_in <- NA
fb$mobile_check_in <- ifelse(fb$mobile_likes>0, "Yes", "NO")
fb$mobile_check_in <- factor(fb$mobile_check_in)
head(fb$mobile_check_in)

#Use Proportion of a Table Function to find out percetage
a<-prop.table(table(fb$mobile_check_in))

paste(a*100,'%')

summary(subset(fb$mobile_check_in, fb$mobile_check_in == "Yes"))


```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!