---
title: "Diamond Analysis"
output: html_notebook
---

```{r}
library(ggplot2)
data(diamonds)

ggplot(aes(x= diamonds$carat, y= diamonds$price), data= diamonds)+
  geom_point(alpha= 0.05)+
  geom_smooth(method= 'lm')+
  geom_abline(slope = 10000, intercept = 0)+
  xlim(c(0,quantile(diamonds$carat,0.99)))+
  ylim(c(0,quantile(diamonds$price,0.99)))
  
```
GGpairs function and sampling the dataset to get and idea of the data in it.
```{r message=FALSE, warning=FALSE, paged.print=TRUE}

set.seed(20022012)
diamonds.samp <- diamonds[sample(1:length(diamonds$price),10000),]

library(scales)
library(GGally)

suppressMessages(ggpairs(diamonds.samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.')))))


```
AS Carat is nothing but volume of the diamonds, lets now look at cuberoot of diamond's carat(x*y*z) is relation to price.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
Math.cbrt <- function(x) {
  sign(x)*abs(x)^(1/3)
}

ggplot(aes(x= Math.cbrt(diamonds$carat), y= diamonds$price), data= diamonds)+
  geom_point(alpha= 0.05)+
  geom_smooth(method= 'lm')+
  xlim(c(0,quantile(diamonds$carat,0.99)))+
  ylim(c(0,quantile(diamonds$price,0.99)))+
  coord_cartesian(xlim = c(0.5:1.5))
```
Check: https://classroom.udacity.com/nanodegrees/nd002/parts/9f3e2782-9409-4c7a-b3d7-6197edfee54e/modules/67c46ff5-5a70-402e-881b-558d374953c6/lessons/755209509/concepts/8675788300923 to get to know when we have to use log scales or not.



```{r}
#Quiz: use grid.arrange

library(gridExtra)

plot1 <- qplot(x=price, data = diamonds, binwidth= 100) + 
  ggtitle('Price')

plot2 <- qplot(x= log10(price), data = diamonds, binwidth= 0.01) +
  ggtitle('Price (log10)')

grid.arrange(plot1,plot2, nrow=1 )
```

Cube Root of Carat to Log 10 Price graph to see the graph of price to carat
```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(scales)

cuberoot_trans = function() trans_new('cuberoot',
  transform= function(x) x^(1/3),
  inverse= function(x) x^3)

p1<- suppressMessages(ggplot(aes(carat, y= price), data= diamonds)+
  geom_point()+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2,3), breaks = c(0.2,0.5,1,2,3))+
  scale_y_continuous(trans= log10_trans(), limits = c(350,15000), breaks = c(350,1000,5000,10000,15000)))+
  ggtitle("Scales changed in this graph")

#Notice the difference between actually changing the data and just changing Axis

p2 <- ggplot(aes(x= Math.cbrt(diamonds$carat), y= log10(diamonds$price)), data= diamonds)+
  geom_point(alpha= 0.05)+
  ggtitle("Values cahnged with Cube Root and Log10")

grid.arrange(p2,p1)

head(diamonds)


```

 Set the transparency to one half, the size to three-fourths, and jitter the points.
 
```{r}
p3= ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.1) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('alpha=0.1 Price (log10) by Cube-Root of Carat')
p4<-ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.5) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('alpha=0.5 Price (log10) by Cube-Root of Carat')

p5<-ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.25) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('alpha=0.25 Price (log10) by Cube-Root of Carat')

p6<-ggplot(aes(carat, price), data = diamonds) + 
  geom_jitter(alpha = 0.05,size = 0.75) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Jitter Price (log10) by Cube-Root of Carat')

suppressWarnings(grid.arrange(p4,p5,p6,p3, ncol=2))
```

Adding more details to carat vs price graph by adding clarity,cut,etc. to the plot
```{r}

library(RColorBrewer)

#Clarity
ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.1, size = 0.5, position = 'jitter', aes(color= clarity)) +
  scale_color_brewer(type = 'div', palette= 'Green',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +
  coord_cartesian(xlim = c(0.2,0.5))+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')

#cut

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.1, size = 1, position = 'jitter', aes(color= cut)) +
  scale_color_brewer(type = 'div', palette= 'RdYlBu',
    guide = guide_legend(title = 'Cut', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +
  coord_cartesian(xlim = c(0.2,0.5))+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')

#Color
ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.1, size = 1, position = 'jitter', aes(color= color)) +
  scale_color_brewer(type = 'div', palette= 'RdYlBu',
    guide = guide_legend(title = 'Color',
    override.aes = list(alpha = 1, size = 2))) +
  coord_cartesian(xlim = c(0.2,0.5))+
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')

  



```
 
 
 Linear modelling of Diamond prices are done below
 
```{r}
library(memisc)

m1<- summary(lm(I(log(price))~I(carat^(1/3)), data= diamonds))
m2<- update(m1, ~. + carat)
m3<- update(m1, ~. + color)
m4<- update(m1, ~. + cut)
m5<- update(m1, ~. + clarity)

?`mtable-class`

mtable(m1,m2,m3,m4, sdigits = 3)

#predictive modelling

thisDiamond = data.frame(carat =1.00, cut= "V.Good",
                         color= "I", clarity="VS1")

modelEstimate = predict(m5, newdata = thisDiamond, 
                        interval= "prediction", level=0.95)

exp(modelEstimate)

```
 
 