Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")

qplot(x= age, y= friend_count, data = pf)

ggplot(aes(age, sqrt(friend_count)), data= pf) + 
  geom_jitter(alpha = 1/50)+ 
  scale_x_continuous(breaks = seq(10,80,10), limits = c(13,80))




```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(age, friend_count), data= pf) + 
  geom_point(alpha = 1/20)+ 
  xlim(13,90)+
  coord_trans(y= 'sqrt')

#adding Jitter to sqrt y cord

ggplot(aes(age, friend_count), data= pf) + 
  geom_point(alpha = 1/20 , position = position_jitter(height = 0))+ 
  xlim(13,90)+
  coord_trans(y= 'sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

ggplot(aes(age, friendships_initiated), data= pf)+
  geom_point(alpha= 1/15, position = position_jitter(height=0))+
  ylim(0,1500)+
  coord_trans(y='sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

library(dplyr)

age_group <- group_by(pf, age)

pf_age_grp = summarise(age_group,fc_mean = mean(friend_count), 
                       fc_median = median(friend_count), n= n())

head(pf_age_grp)

#chaining functions

pf_age_grp_c <- pf %>%
  group_by(age)%>%
  summarise(fc_mean = mean(friend_count),
            fc_meadian = median(friend_count),fc_max = max(friend_count), n= n())%>%
  arrange(age)

pf_age_in_mon_grp_c <- pf %>%
  group_by(age_in_months)%>%
  summarise(fc_mean = mean(friend_count),
            fc_meadian = median(friend_count),fc_max = max(friend_count), n= n())%>%
  arrange(age_in_months)
?summarise

head(pf_age_in_mon_grp_c,10)

```

Create your plot!

```{r fig.width= 100, fig.asp= 0.65 Conditional Means Plot}

ggplot(aes(x= age, y= fc_mean), data= pf_age_grp_c)+
  geom_point()+
  scale_x_continuous(breaks = seq(10,80,5), limits = c(10,80))+
  ylim(50,400)

ggplot(aes(x= age, y= fc_mean), data= pf_age_grp_c)+
  geom_line()+
  scale_x_continuous(breaks = seq(10,80,5), limits = c(10,80))+
  ylim(50,400)

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(age, friend_count), data= pf)+
  geom_point(alpha= 0.05, position = position_jitter(height=0), color = 'orange')+
  coord_trans(y='sqrt')+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = quantile, fun.args= list(probs = .25), linetype = 5,size = 1, color= 'white')+
  geom_line(stat = 'summary', fun.y = quantile, fun.args= list(probs = .9), linetype = 2, color= 'blue')+
  coord_cartesian(xlim = c(13,90), ylim = c(0,1000))
  

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age,pf$friend_count, method = c("pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
#Whole Dataset
with(data= pf, cor.test(age, friend_count, method = c("pearson")))

#subsetof dataset this is Non-parametric test not generally used
with(data= subset(pf,pf$age <= 70), cor.test(age, friend_count, method = c("kendall")))

#subsetof dataset used for Ordinal data (lik=1,somewhat like =2, neutral =3,etc.)
with(data= subset(pf,pf$age <= 70), cor.test(age, friend_count, method = c("spearman")))

#subsetof dataset most commly used to denote correlation
with(data= subset(pf,pf$age <= 70), cor.test(age, friend_count, method = c("pearson")))

```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

ggplot(aes(x= www_likes_received, y= likes_received), data= pf)+
  geom_point()+
  coord_cartesian(xlim = c(0,quantile(pf$www_likes_received,0.95)),ylim = c(0,quantile(pf$likes_received,0.95)))+
  geom_smooth(method = 'lm', color='white', linetype =5)

cor.test(pf$www_likes_received,pf$likes_received)

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received,pf$likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data("Mitchell")

ggplot(aes(x=Mitchell$Month,y=Mitchell$Temp), data= Mitchell)+
  geom_point()

str(Mitchell)

table(Mitchell$month,Mitchell$Temp)
  

```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
No strong correlation, but this seems like a trick question, for now I will play along.
b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Temp,Mitchell$Month)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

ggplot(aes(x=((Mitchell$Month%%12)+1),y=Mitchell$Temp), data= Mitchell)+
  geom_point()+
  scale_x_continuous(breaks = seq(1,12,1))

tail(Mitchell)

ggplot(aes(x=Mitchell$Month,y=Mitchell$Temp), data= Mitchell)+
  geom_point()+
  scale_x_continuous(breaks = seq(0,203,12))

str(Mitchell)

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

pf$age_in_months <- pf$age + (1-pf$dob_month/12)

tail(pf$age_in_months)

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

p1=ggplot(aes(x= age_in_months, y= fc_mean), data= pf_age_in_mon_grp_c)+
  geom_line()+
  scale_x_continuous(breaks = seq(10,70,5), limits = c(10,70))+
  ylim(50,400)+
  geom_smooth()

p2=ggplot(aes(x= round((age/5))*5, y= friend_count), data= pf)+
  geom_line(stat = 'summary', fun.y= mean)+
  scale_x_continuous(breaks = seq(10,70,5), limits = c(10,70))+
  ylim(50,400)

library(gridExtra)

grid.arrange(p1,p2,ncol=1)

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

