Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library(dplyr)
library(ggplot2)

pf <- read.csv('pseudo_facebook.tsv', sep= '\t')
head(pf)

pf.fc_by_age_gender = pf%>%
  filter(!is.na(gender))%>%
  dplyr::group_by(age,gender)%>%
  dplyr::summarise(mean_friend_count = mean(as.numeric(friend_count)), 
                       median_friend_count = as.integer(median(as.numeric(friend_count))), n= n())%>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)
tail(pf.fc_by_age_gender)

ggplot(aes(x= age, y= median_friend_count),
       data= pf.fc_by_age_gender)+
geom_line(aes(color= gender))
  
```

***

### Thinking in Ratios
Notes:



***

### Wide and Long Format
Notes:

Convert from Long

Age Friend_Count  Gender
13  Male          X1
13  Female        ...
14  Male
14  Female

To Wide

Age Male  Female
13  X1    X2
14  ...
***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

#this is in wide format
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age~gender, 
                                  value.var = 'median_friend_count')

#to re-eshtablish long format
tail(melt(pf.fc_by_age_gender.wide, id.vars = c("age")))


#tail(pf.fc_by_age_gender.wide)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x=age , y= female/male), data= pf.fc_by_age_gender.wide)+
  geom_line()+
  geom_hline(yintercept = 1, linetype=2,alpha =0.3)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

#year_joined varibale creation

head(pf$tenure)

pf$year_joined = (2014 - ceiling((pf$tenure/365)))

table(pf$year_joined.bucket)

pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009,2011,2012,2014))

?cut

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}

ggplot(aes(x= age, y= friend_count),
       data= subset(pf,!(is.na(year_joined.bucket))))+
geom_line(aes(color= year_joined.bucket), stat = 'summary', fun.y= median)

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
#adding grandmean in

ggplot(aes(x= age, y= friend_count),
       data= subset(pf,!(is.na(year_joined.bucket))))+
geom_line(aes(color= year_joined.bucket), stat = 'summary', fun.y= mean)+
  geom_line(stat='summary',fun.y=mean,linetype=2, color='black')

pf.subset_min_1_day <- subset(pf,pf$tenure>=1)

#summary with tenure
summary(pf.subset_min_1_day$friend_count/pf.subset_min_1_day$tenure)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}
ggplot(aes(x= 20*round(tenure/20), y= friendships_initiated/tenure),
       data= subset(pf,pf$tenure>=1))+
geom_line(aes(color= year_joined.bucket), stat = 'summary', fun.y= mean)

#Same graph with a smoother

ggplot(aes(x= tenure, y= friendships_initiated/tenure),
       data= subset(pf,pf$tenure>=1))+
geom_smooth(aes(color= year_joined.bucket))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

yo <- read.csv('yogurt.csv')

yo$id <- factor(yo$id)
str(yo)

head(yo)

summary(yo)

ggplot(aes(price), data = yo)+
  geom_histogram(binwidth = 10)+
  geom_vline(xintercept = mean(yo$price), color= 'red', linetype=2)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}


yo <- transform(yo,all.purchases = (strawberry+blueberry+
                                pina.colada+plain+mixed.berry))

head(yo)

```

***

### Prices over Time
Notes:

```{r Prices over Time}

ggplot(aes(time/365,price), data=yo)+
  geom_jitter(alpha= 0.2, shape= 21)

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r message=FALSE,warning=FALSE Looking at Sample of Households}

set.seed(400)
?levels
sample.ids <- sample(yo$id,16)

sample.ids

ggplot(aes(time,price), data=subset(yo, id %in% sample.ids))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size = all.purchases), pch= 1)


#multiple scatter plots put together

library(GGally)

theme_set(theme_minimal(20))

#producing multiple random graphs

set.seed(1836)
pf.subset <- pf[,c(2:15)]
names(pf.subset)
suppressWarnings(ggpairs)
ggpairs(pf.subset[sample.int(nrow(pf.subset),1000),])

pf1 <- pf[sample.int(nrow(pf), 1000), -1]

cor.test(pf1$friendships_initiated,pf1$friend_count)
cor.test(pf1$mobile_likes,pf1$age)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

